\speculake: Hiding Malware in Speculative Execution

### 2018-05-09 paper/discussion.tex
typos

Finally, \speculake could also use another side channel method besides that
avoids the cache
to exfiltrate information from the speculative gadgets, such as the branch
predictor itself~\cite{evtyushkin2018branchscope}, memory bandwidth, or power
utilization of the system.

@@

\texttt{Guest1 -> Guest1} - When both trigger and payload are run internal to
a single virtual machine the branch prediction is successfully trained in the payload
program, though at a notably reduced hit rate. This can be overcome by using an 
increased number of iterations to establish confidence.

\texttt{Guest -> Host} - In a configuration where trigger runs on the guest while the 
payload runs on the host we have found that \speculake also successfully trains the 
branch predictor well enough retreive a signal. Again, this signal is significantly
reduced, implying the hypercall context swtich from guest to host requires some 
overhead. We note that processor affinity in a guest is not mapped to processor 
affinity in the host, as the host assigns the vm to any available CPU, though 
eventually the hypervisor will assign the vm back to the same core as the payload 
program allowing progress to continue.

\texttt{Host -> Guest} - When trigger runs on the host attempting to train the 
branch predictor to influence a payload process in a guest we have found that 
the payload receives no signal. This implies that the hypercall context switch 
from the host to the guest is heavier and has a larger impact on branch history. 
This matches with work done by the Project Zero team at Google which found that 
hypervisors can be fingerprinted from within a vm using the branch 
predictor~\cite{project_zero}.

\texttt{Guest1 -> Guest2} - Configurations where trigger and payload 
run on coresident virtual machines similarly show no signal for the same 
reasons that host to guest \speculake fails. However, this has the added difficulty
that virtual machines must be run on the same core or on a pair of 
hyperthreaded cores.

@@

Alternatively multiple triggers could be used make individual state changes in
the payload program, and write in data from a local, or (as demonstrated in
section~\ref{subsec:openssl}) remote processes. For example, a payload program
can check a series of three distinct jump patterns corresponding to three
different triggers. If the first trigger is detected the payload shifts in a 1
to the state; if the second is detected it shifts in a 0. If the third trigger
is detected it discontinues listening to the three triggers and executes the
data shifted into the state.
This could be a direct form of \speculake using emulation and
decryption or more traditional malicious behavior. Using this method a payload
program can be distributed without the encrypted data section all together,
instead having the data written into the process at run-time without ever
connecting out to any other resource.

\subsubsection{Benign Triggers}
We have demonstrated that a using OpenSSL as a benign trigger allows for
\speculake payloads to be triggered remotely, and discussed the minimum
requirements for identifying a benign trigger. However, we have not classified
typical system resources that might be used as benign triggers. This is a
promising direction for future investigation as various system libraries present
juicy targets for trigger design. For exmaple, if a sufficient series of jumps
is found to exist in the TCP stack implemented in the unix kernel (i.e. in the
TCP Checksum library) an attacker could reasonably trigger a \speculake payload
remotely, from a spoofed IP address. This attack would rely on a payload
defeating KASLR, however, this has been demonstrated to be
feasible~\cite{gruss2017kaslr,evtyushkin2016jump}

>>>>

Finally, \speculake could also use another side channel method that
avoids the cache
to exfiltrate information from the speculative gadgets, such as the branch
predictor itself~\cite{evtyushkin2018branchscope}, memory bandwidth, or power
utilization of the system.

@@

\texttt{Guest1 -> Guest1}---When both trigger and payload are run internal to a
single virtual machine the branch prediction is successfully trained in the
payload program, though at a notably reduced hit rate. This can be overcome by
using an increased number of iterations to establish confidence.

\texttt{Guest -> Host}---In a configuration where trigger runs on the guest
while the payload runs on the host we have found that \speculake also
successfully trains the branch predictor well enough retreive a signal. Again,
this signal is significantly reduced, implying the hypercall context swtich from
guest to host requires some overhead. We note that processor affinity in a guest
is not mapped to processor affinity in the host, as the host assigns the VM to
any available CPU, though eventually the hypervisor will assign the VM back to
the same core as the payload program allowing progress to continue.

\texttt{Host -> Guest}---When trigger runs on the host attempting to train the
branch predictor to influence a payload process in a guest we have found that
the payload receives no signal. This implies that the hypercall context switch
from the host to the guest is heavier and has a larger impact on branch history.
This matches with work done by the Project Zero team at Google which found that
hypervisors can be fingerprinted from within a VM using the branch
predictor~\cite{project_zero}.

\texttt{Guest1 -> Guest2}---Configurations where trigger and payload run on
co-resident virtual machines similarly show no signal for the same reasons that
host to guest \speculake fails. However, this has the added difficulty that
virtual machines must be run on the same core or on a pair of hyperthreaded
cores.

@@

Alternatively multiple triggers could be used make individual state changes in
the payload program, and write in data from a local, or (as demonstrated in
Section~\ref{subsec:openssl}) remote processes. For example, a payload program
can check a series of three distinct jump patterns corresponding to three
different triggers. If the first trigger is detected the payload shifts in a 1
to the state; if the second is detected it shifts in a 0. If the third trigger
is detected it discontinues listening to the three triggers and executes the
data shifted into the state.
This could be a direct form of \speculake using emulation and
decryption or more traditional malicious behavior. Using this method a payload
program can be distributed without the encrypted data section all together,
instead having the data written into the process at run-time without ever
connecting out to any other resource.

\subsubsection{Benign Triggers}
We have demonstrated that a using \allowbreak OpenSSL as a benign trigger allows for
\speculake payloads to be triggered remotely, and discussed the minimum
requirements for identifying a benign trigger. However, we have not classified
typical system resources that might be used as benign triggers. This is a
promising direction for future investigation as various system libraries present
juicy targets for trigger design. For exmaple, if a sufficient series of jumps
is found to exist in the TCP stack implemented in the UNIX kernel (i.e. in the
TCP Checksum library) an attacker could reasonably trigger a \speculake payload
remotely, from a spoofed IP address. This attack would rely on a payload
defeating KASLR, however, this has been demonstrated to be
feasible~\cite{gruss2017kaslr,evtyushkin2016jump}.
